---
title: Fitness Tracker data analysis
output: html_document
---
  
## Introduction and background
This is a simple analysis performed on data from fitness trackers

#### Upload CSV files to R
Files used for analysis uploaded from kaggle:
https://www.kaggle.com/arashnic/fitbit


#### Installing and loading common packages and libraries

```{r}
library(tidyverse)
library(readr)
library(writexl)
```


#### Loading CSV files
creating a dataframe  called "daily_activity" to upload specific data for analysis from "Fitabase Data 4.12.16-5.12.16".

```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
```
creating another dataframe called "sleep_day" to upload more data 
```{r}
sleep_day <- read.csv("sleepDay_merged.csv")
```

#### Exploring data

Taking a look at the daily_activity data.
```{r}
head(daily_activity)
```

Identifying all the columns in the daily_activity data.
```{r}
colnames(daily_activity)
```

Taking a look at the sleep_day data.
```{r}
head(sleep_day)
```

Identifying all the columns in the daily_activity data.
```{r}
colnames(sleep_day)
```

both datasets have the 'Id' field - this can be used to merge the datasets.

#### Understanding some summary statistics
first, identifying the unique participants in each dataframe

```{r distinct users}
n_distinct(daily_activity$Id)
n_distinct(sleep_day$Id)
```

next how many observations are there in each dataframe?
```{r observations}
nrow(daily_activity)
nrow(sleep_day)
```

#### More detailed statistics summary
  
  For the daily activity dataframe:
```{r}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()
```
  

For the sleep dataframe:
```{r}
sleep_day %>%  
  select(TotalSleepRecords,
         TotalMinutesAsleep,
         TotalTimeInBed) %>%
  summary()
```
from the summary, the participants walked an average of 7638 steps daily and slept for 7 hours on average

#### Plotting a few explorations

Now to look at the relationship between steps taken in a day and sedentary minutes. This could help inform us on customer market segments to market to.  

```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryMinutes)) + geom_point()+ 
  geom_smooth()+
  labs(title = "Total steps vs Sedentary Minutes", x = "Total steps", y = "Sedentary Minutes")
```

Next to look at the relationship between minutes asleep and time in bed. 
  
```{r}
ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + geom_point()+ 
  geom_smooth()+
  labs(title = "Total Minutes Asleep vs Total time in bed", x = "Total minutes Asleep", y = "Total time in bed")
```
  

  
#### Merging these two datasets together
  
```{r}
combined_data <- merge(sleep_day, daily_activity, by="Id")
head(combined_data)
```
#### A summary of combined_data

```{r}
head(combined_data)
```

```{r}
colnames(combined_data)
```

```{r}
combined_data %>%
  select(TotalSteps, TotalDistance, Calories, SedentaryMinutes, VeryActiveMinutes, TotalTimeInBed, TotalMinutesAsleep) %>%
  summary()
```


Taking a look at how many participants are in this data set.

```{r}
n_distinct(combined_data$Id)
```
# Plotting a few more Explorations

Looking at the relationship between daily steps and daily calories
```{r}
ggplot(combined_data, aes(x = TotalSteps, y = Calories))+
  geom_point()+
  geom_smooth()+
  labs(title = "Daily steps vs Calories", x = "Daily steps", y = "Calories")
```

Now a look at the relationship between daily exercise time(VeryActiveMinutes) and calories
```{r}
ggplot(combined_data, aes(x = VeryActiveMinutes, y = Calories))+
  geom_point()+
  geom_smooth()+
  labs(title = "Daily exercise vs Calories", x = "Active Minutes", y = "Calories")
```

Now a look at the relationship between daily sleep and calories
```{r}
ggplot(combined_data, aes(x = TotalMinutesAsleep, y = Calories))+
  geom_point()+
  geom_smooth()+
  labs(title = "Daily sleep vs Calories", x = "Daily sleep time(minutes)", y = "Calories")
```


 Exporting dataframe for further visualization on Tableau

```{r}
write_xlsx(x = combined_data, path = "combined_data.xlsx", col_names = TRUE)
```

